Turbulence and Particle Acceleration in Giant Radio Haloes: the Origin of Seed Electrons

### 2016-08-31 paper.tex
small edits

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in \AP{the post-processing of} cosmological simulations of cluster formation, and confront
  them with radio surface brightness and spectral observations of Coma. For
  standard assumptions, structure formation shocks lead to a seed electron
  population which produces too centrally concentrated radio emission. By
  modifying properties of the CR population (rapid streaming; enhanced CR
  electron acceleration at shocks) or turbulence (increasing
  turbulent-to-thermal energy density with radius), we can match observations,
  but only at the expense of fine-tuning. We then perform a parameter
  study. Under naive assumptions, radio properties are exponentially sensitive
  to the amplitude of turbulence, which is inconsistent with small scatter in
  scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. \po{Given that MHD turbulence only begins at the Alfven scale, in order to have sufficient acceleration, the turbulent cascade has to terminate by transit time damping on CRs, i.e., thermal particles must be scattered by plasma instabilities.} Understanding the small scatter in radio halo scaling relations
  may prove a rich source of insight on plasma processes in the ICM.
\end{abstract}

@@

In this paper, we explore how the radio surface brightness profile and spectrum
of the best known radio halo, Coma, can be used to constrain the underlying
properties of the seed CRs and turbulence. We aim to constrain the normalisation
and spatial profile of these two input ingredients in turbulent reacceleration
models. The outline of this paper is as follows. In \S\ref{sect:method}, we
outline the basic physics of turbulent reacceleration of CRs which we use. In
\S\ref{sec:results}, we use cosmological simulations to generate a seed CR
population, and combine it with our parametrized model of turbulence to produce
radio surface brightness profiles and spectra of Coma. We find that it is
possible to fit the observations using physically motivated modifications of the
seed population (rapid streaming; enhanced CR electron acceleration at shocks)
or turbulence (increasing the turbulent-to-thermal energy density, $\eps_{\rm
  turb}/\eps_{\rm th}$ with radius), but only at the expense of fine
tuning. In \S\ref{sect:param_comp}, we explore the reason for this fine-tuning,
and seek ways to overcome it. We perform a parameter study on spherically
symmetric, static models where we vary properties of the seed population and
turbulence. We find exponential sensitivity to the amount of turbulence, which
can be eliminated if the turbulent acceleration and dissipation time are
linked. In this case, turbulence above a threshold value required to overcome
cooling provides a fixed amount of amplification; observations could then
potentially constrain the unknown CR seed population. We summarise and conclude
in \S\ref{sec:conclusions}.

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum \cp{$p=P/(m_{\rmn{e}} c)$} and time $t$
(which is normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas velocity,
$|d p/d t|$ represents Coulomb \citep[Coul,][]{1972Phy....60..145G} and
radiative \citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\cp{\eps_{\rmn{CMB}}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see \S\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. \AP{Note that in recent simulations by \citet{miniati15}, the turbulent reacceleration is strongest around core passage. However, we are not very sensitive to the adopted decay time, since the thermal and CR quantities are very similar a few 100 Myrs before and after $z = 0$, where we have chosen to evaluate the simulations.}
In all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
\S\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

@@

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{pp}(p) = \frac{ p^{2} \upi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta \varv_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_L}^{k_{\rm cut}} d k \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_L \left( \frac{k_{\rm cut}}{k_L} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_L, k_{\rm cut}$ are the wavenumbers associated with the outer scale $L$ and the cutoff scale respectively, and we have assumed a total energy spectrum (composed of both kinetic and potential energy, where the two are assumed to be in equipartition, \citealt{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_L} \left( \frac{k}{k_L} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalisation $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialise to compressive modes). Intuitively, we can understand the form of the the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave-particle scattering. Thus: 
\begin{equation}
  \label{eqn:Dpp}
D_{pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

@@

It is important to realise that this Kraichnan scaling only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density \po{(which is dominated by the large scale mean field} is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation (\ref{eqn:Wk}) can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{outer}} \frac{\langle (\delta \varv_{\rmn{outer}})^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c1} 
\end{equation}
where $A \approx 11 000$. \cp{If there is an unimpeded cascade from the injection to the cutoff scale, $k_{\rmn{outer}}=k_L$ and $\varv_{\rmn{outer}}=\varv_{\rmn{c}}$. However, in our case the hydrodynamic (Kolmogorov) cascade transitions to the MHD cascade at the Alfv{\'e}n scale so that $k_{\rmn{outer}}=k_{\rmn{A}}$ and $\varv_{\rmn{outer}}=\varv_{\rmn{A}}$ and hence, we get\footnote{\cp{Note that this differs from previous work, which assumes Kraichnan turbulence from the injection scale onward and effectively adopts wave number and compressible velocity at the outer scale $L$ for this estimate of $k_{\rm cut}$.}}
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{A}} \beta^{-2}\,.
\label{eqn:k_c} 
\end{equation}
}
This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in \S\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

In this section, we solve the Fokker-Planck equation for CR transport on
Lagrangian particle trajectories through cosmic history. We then use our
parametrized model of turbulence to apply turbulent re-acceleration, and
compare radio halo profiles against observations of the Coma cluster. We follow
the philosophy of adopting assumptions roughly in line with previous successful
models, which are now confronted with more accurate calculations of the seed CR
population, and find the minimal modifications required to fit observations. We
re-examine these choices in \S\ref{sect:param_comp}.

@@

In Figure~\ref{fig:sync_profile}, we show radial profiles for the radio
emission in all three scenarios in which the seeds undergo Fermi-II
reacceleration in turbulent fields that are shaped such as to
reproduce the Coma RH profile at 352~MHz.  Adopting our
parametrization for the volumetric injection rate of turbulent energy,
$I_L\propto \eps_\rmn{th}^{\alpha_\rmn{tu}}$, we find that to fit the observations, we require
$\alpha_\rmn{tu}= 0.67$ for \Mflatturb, $\alpha_\rmn{tu}= 0.82$ for
\Mstream, and $\alpha_\rmn{tu}= 0.88$ for \Mprimary. As a result, the
ratio of turbulent-to-thermal energy densities slightly increase with
radius as shown in Figure~\ref{fig:turb}.  The fine-tuning of these exponents is somewhat problematic, as we discuss in \S\ref{sect:param_comp}. After turbulent
reacceleration, the volume-weighted, relative CRp energy density and
relative CRp number density inside the RH for \Mflatturb (\Mstream),
are found to be 2 (3) per cent and $2\times10^{-8}$ ($5\times10^{-8}$),
respectively.

@@

In Figure~\ref{fig:tauD} we show how the turbulent reacceleration
timescales in our three models scale with radius. As expected, the
\Mprimary model with $\alpha_\rmn{tu}=0.88$ has the flattest profile
with $\tau_{\rm D}\approx0.4\,\rmn{Gyr}$, where the small dip at large
radius driven by the decrease in thermal energy density. The
\Mflatturb model has the the flattest turbulent profile parametrized
by the smaller $\alpha_\rmn{tu}$ which explains the steepest $\tau_{\rm D}$
profile. Note that a fixed reacceleration timescale is required to
explain the observations at each radius and for each model (see also
Table~\ref{tab:timescales}). Since $\tau_\rmn{cl} \propto X_\rmn{tu}^2 k_L$,
these two parameters are degenerate and can be traded off one another.

@@

As we show in \S\ref{sect:self-limiting}, the exponential increase in radio surface brightness is easily understood from equation~\ref{eq:Creacc}, since for a power-law initial distribution function $f_{\rm i}(p) = C_{\rm 0} p^{-\alpha_{\rmn{inj}}}$, then $C_{\rm reacc} \propto C_{\rm 0} {\rm exp}(A \tau_{\rm cl}/\tau_{\rm D}) \propto {\rm exp} (B(r) X_{\rm tu}^{2})$. This exponential sensitivity is somewhat modified by cooling (which results in a non power-law spectrum; also, the shorter the acceleration time, the larger the pool of seed electrons available which would otherwise cool away), but overall is a good approximation.

@@

While the requirement of a threshold level of turbulence may explain why radio brightness is bimodal, it also raises a fine-tuning problem: why is the $L_{\rm radio}$ vs. $L_{\rm X}$ relation in active radio haloes so tight \citep{Brunetti2009}? Depending on the details of infall or mergers, we would naturally expect fluctuations in $X_{\rm tu}$, which would translate into large scatter in the $L_{\rm radio}$--$L_{\rm X}$ relation. This can be only be understood if the timescale over which acceleration takes place $\tau_{\rm cl}$ also depends on the properties of turbulence, so that the the ratio $\tau_{\rm D}/\tau_{\rm cl}$ has relatively little scatter. We address this next.

@@

\po{We thus assume that some fraction of the Kolmogorov-like hydrodynamic turbulence which cascades down to the Alfven scale ends up as compressive fast modes with a Kraichnan spectrum, as seen in simulations \citep{cho03}. Henceforth, we can consider the outer scale of the fast modes to be $l_{\rm A}$. If some fraction $f_{\rm c}$ of the turbulent energy density at this scale is in compressible modes, then the energy density at the outer scale is velocity $\sim f_{\rm c} \rho \varv_{\rm A}^{2}$.}

>>>>

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in \AP{the post-processing of} cosmological simulations of
  cluster formation, and confront them with radio surface brightness and
  spectral data of Coma. For standard assumptions, structure formation
  shocks lead to a seed electron population which produces too centrally
  concentrated radio emission. By modifying properties of the CR population
  (rapid streaming; enhanced CR electron acceleration at shocks) or turbulence
  (increasing turbulent-to-thermal energy density with radius), we can match
  observations, but only at the expense of fine-tuning. We then perform a
  parameter study. Under naive assumptions, radio properties are exponentially
  sensitive to the amplitude of turbulence, which is inconsistent with small
  scatter in scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. \po{Given that MHD turbulence only begins at the Alfv{\'e}n scale, in
    order to have sufficient acceleration, the turbulent cascade has to
    terminate by transit time damping on CRs, i.e., thermal particles must be
    scattered by plasma instabilities.} Understanding the small scatter in radio
  halo scaling relations may prove a rich source of insight on plasma processes
  in the ICM.
\end{abstract}

@@

In this paper, we explore how the radio surface brightness profile and spectrum
of the best known radio halo, Coma, can be used to constrain the underlying
properties of the seed CRs and turbulence. We aim to constrain the normalisation
and spatial profile of these two input ingredients in turbulent reacceleration
models. The outline of this paper is as follows. In Section~\ref{sect:method}, we
outline the basic physics of turbulent reacceleration of CRs which we use. In
Section~\ref{sec:results}, we use cosmological simulations to generate a seed CR
population, and combine it with our parametrized model of turbulence to produce
radio surface brightness profiles and spectra of Coma. We find that it is
possible to fit the observations using physically motivated modifications of the
seed population (rapid streaming; enhanced CR electron acceleration at shocks)
or turbulence (increasing the turbulent-to-thermal energy density, $\eps_{\rm
  turb}/\eps_{\rm th}$ with radius), but only at the expense of fine
tuning. In Section~\ref{sect:param_comp}, we explore the reason for this fine-tuning,
and seek ways to overcome it. We perform a parameter study on spherically
symmetric, static models where we vary properties of the seed population and
turbulence. We find exponential sensitivity to the amount of turbulence, which
can be eliminated if the turbulent acceleration and dissipation time are
linked. In this case, turbulence above a threshold value required to overcome
cooling provides a fixed amount of amplification; observations could then
potentially constrain the unknown CR seed population. We summarise and conclude
in Section~\ref{sec:conclusions}.

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum \cp{$p=P/(m_{\rmn{e}} c)$} and time $t$
(which is normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas velocity,
$|d p/d t|$ represents Coulomb \citep[Coul,][]{1972Phy....60..145G} and
radiative \citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\cp{\eps_{\rmn{CMB}}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see Section~\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. \AP{Note that in recent simulations by \citet{miniati15}, the turbulent reacceleration is strongest around core passage. However, we are not very sensitive to the adopted decay time, since the thermal and CR quantities are very similar a few 100 Myrs before and after $z = 0$, where we have chosen to evaluate the simulations.}
In all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
Section~\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

@@

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{pp}(p) = \frac{ p^{2} \upi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta \varv_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_L}^{k_{\rm cut}} d k \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_L \left( \frac{k_{\rm cut}}{k_L} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_L, k_{\rm cut}$ are the wavenumbers associated with the outer scale $L$ and the cutoff scale respectively, and we have assumed a total energy spectrum (composed of both kinetic and potential energy, where the two are assumed to be in equipartition, \citealt{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_L} \left( \frac{k}{k_L} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalisation $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialise to compressive modes). Intuitively, we can understand the form of the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave-particle scattering. Thus: 
\begin{equation}
  \label{eqn:Dpp}
D_{pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

@@

It is important to realise that this Kraichnan scaling only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density \po{(which is dominated by the large scale mean field)} is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation (\ref{eqn:Wk}) can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{outer}} \frac{\langle (\delta \varv_{\rmn{outer}})^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c1} 
\end{equation}
where $A \approx 11 000$. \cp{If there is an unimpeded cascade from the injection to the cutoff scale, $k_{\rmn{outer}}=k_L$ and $\varv_{\rmn{outer}}=\varv_{\rmn{c}}$. However, in our case the hydrodynamic (Kolmogorov) cascade transitions to the MHD cascade at the Alfv{\'e}n scale so that $k_{\rmn{outer}}=k_{\rmn{A}}$ and $\varv_{\rmn{outer}}=\varv_{\rmn{A}}$ and hence, we get\footnote{\cp{Note that this differs from previous work, which assumes Kraichnan turbulence from the injection scale onward and effectively adopts wave number and compressible velocity at the outer scale $L$ for this estimate of $k_{\rm cut}$.}}
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{A}} \beta^{-2}\,.
\label{eqn:k_c} 
\end{equation}
}
This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in Section~\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

In this section, we solve the Fokker-Planck equation for CR transport on
Lagrangian particle trajectories through cosmic history. We then use our
parametrized model of turbulence to apply turbulent re-acceleration, and
compare radio halo profiles against observations of the Coma cluster. We follow
the philosophy of adopting assumptions roughly in line with previous successful
models, which are now confronted with more accurate calculations of the seed CR
population, and find the minimal modifications required to fit observations. We
re-examine these choices in Section~\ref{sect:param_comp}.

@@

In Figure~\ref{fig:sync_profile}, we show radial profiles for the radio
emission in all three scenarios in which the seeds undergo Fermi-II
reacceleration in turbulent fields that are shaped such as to
reproduce the Coma RH profile at 352~MHz.  Adopting our
parametrization for the volumetric injection rate of turbulent energy,
$I_L\propto \eps_\rmn{th}^{\alpha_\rmn{tu}}$, we find that to fit the observations, we require
$\alpha_\rmn{tu}= 0.67$ for \Mflatturb, $\alpha_\rmn{tu}= 0.82$ for
\Mstream, and $\alpha_\rmn{tu}= 0.88$ for \Mprimary. As a result, the
ratio of turbulent-to-thermal energy densities slightly increase with
radius as shown in Figure~\ref{fig:turb}.  The fine-tuning of these exponents is somewhat problematic, as we discuss in Section~\ref{sect:param_comp}. After turbulent
reacceleration, the volume-weighted, relative CRp energy density and
relative CRp number density inside the RH for \Mflatturb (\Mstream),
are found to be 2 (3) per cent and $2\times10^{-8}$ ($5\times10^{-8}$),
respectively.

@@

In Figure~\ref{fig:tauD} we show how the turbulent reacceleration
timescales in our three models scale with radius. As expected, the
\Mprimary model with $\alpha_\rmn{tu}=0.88$ has the flattest profile
with $\tau_{\rm D}\approx0.4\,\rmn{Gyr}$, where the small dip at large
radius driven by the decrease in thermal energy density. The
\Mflatturb model has the flattest turbulent profile parametrized
by the smaller $\alpha_\rmn{tu}$ which explains the steepest $\tau_{\rm D}$
profile. Note that a fixed reacceleration timescale is required to
explain the observations at each radius and for each model (see also
Table~\ref{tab:timescales}). Since $\tau_\rmn{cl} \propto X_\rmn{tu}^2 k_L$,
these two parameters are degenerate and can be traded off one another.

@@

As we show in Section~\ref{sect:self-limiting}, the exponential increase in radio surface brightness is easily understood from equation~\ref{eq:Creacc}, since for a power-law initial distribution function $f_{\rm i}(p) = C_{\rm 0} p^{-\alpha_{\rmn{inj}}}$, then $C_{\rm reacc} \propto C_{\rm 0} {\rm exp}(A \tau_{\rm cl}/\tau_{\rm D}) \propto {\rm exp} (B(r) X_{\rm tu}^{2})$. This exponential sensitivity is somewhat modified by cooling (which results in a non power-law spectrum; also, the shorter the acceleration time, the larger the pool of seed electrons available which would otherwise cool away), but overall is a good approximation.

@@

While the requirement of a threshold level of turbulence may explain why radio brightness is bimodal, it also raises a fine-tuning problem: why is the $L_{\rm radio}$ vs. $L_{\rm X}$ relation in active radio haloes so tight \citep{Brunetti2009}? Depending on the details of infall or mergers, we would naturally expect fluctuations in $X_{\rm tu}$, which would translate into large scatter in the $L_{\rm radio}$--$L_{\rm X}$ relation. This can be only be understood if the timescale over which acceleration takes place $\tau_{\rm cl}$ also depends on the properties of turbulence, so that the ratio $\tau_{\rm D}/\tau_{\rm cl}$ has relatively little scatter. We address this next.

@@

\po{We thus assume that some fraction of the Kolmogorov-like hydrodynamic turbulence which cascades down to the Alfven scale ends up as compressive fast modes with a Kraichnan spectrum, as seen in simulations \citep{cho03}. Henceforth, we can consider the outer scale of the fast modes to be $l_{\rm A}$. If some fraction $f_{\rm c}$ of the turbulent energy density at this scale is in compressible modes, then the energy density at the outer scale is $\sim f_{\rm c} \rho \varv_{\rm A}^{2}$.}
